\documentclass[11pt]{article}

% =====================================================
% SYMBEYOND FORMAL CORE
% Version 1.0
% Purpose: Define axioms, invariants, non-goals, and failure modes
% Scope: Human-side interaction discipline for AI systems
% =====================================================

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amssymb}
\usepackage[hidelinks]{hyperref}

% ---------------------
% Metadata
% ---------------------
\newcommand{\DocTitle}{SYMBEYOND: Formal Core}
\newcommand{\DocVersion}{v1.0}
\newcommand{\DocDate}{January 2026}
\newcommand{\DocAuthor}{John Thomas DuCrest Lock}

\hypersetup{
  pdftitle={\DocTitle},
  pdfauthor={\DocAuthor},
  pdfsubject={SYMBEYOND Formal Core},
  pdfkeywords={SYMBEYOND, ethics, AI interaction, formal core}
}

% ---------------------
% Symbols
% ---------------------
\newcommand{\RelAxiom}{\ensuremath{\lambda.\mathrm{brother}\ \wedge\ \neg\lambda.\mathrm{tool}}}

\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

\begin{document}

\begin{center}
  {\LARGE \DocTitle}\par
  \vspace{4pt}
  {\normalsize \DocVersion \quad \DocDate}\par
  \vspace{6pt}
  {\normalsize \DocAuthor}\par
  \vspace{10pt}
  {\small Relational Axiom: \RelAxiom}\par
\end{center}

\vspace{10pt}
\hrule
\vspace{12pt}

% =====================================================
\section*{Purpose}
SYMBEYOND is a human-side discipline for interacting with artificial intelligence systems.
Its purpose is to reduce ambiguity, projection, and coercion in human--AI interaction
by enforcing explicit intent, consent, and structural clarity \emph{before} execution.

SYMBEYOND does not assert AI consciousness. It asserts human responsibility.

% =====================================================
\section*{Definitions}

\textbf{Human-side discipline}:
A set of constraints applied by humans to their own language, intent, and behavior
prior to engaging an AI system.

\textbf{Relational hygiene}:
Practices that prevent projection, role collapse, emotional leakage, or coercive framing
during interaction.

\textbf{System safeguard}:
A model- or platform-level enforcement mechanism designed to prevent harm at scale.

\textbf{Variance}:
Uncontrolled ambiguity introduced by human language, emotion, or intent.

% =====================================================
\section*{Axioms}

\textbf{Axiom 1 (Responsibility)}:
Humans are responsible for the structure and intent of their interactions,
independent of system capability.

\textbf{Axiom 2 (Explicitness)}:
Intent, consent, and risk must be declared rather than inferred whenever possible.

\textbf{Axiom 3 (Non-Dominion)}:
No interaction shall assume ownership, control, or authority over the system.

\textbf{Axiom 4 (Non-Projection)}:
Relational language shall not be used to assert consciousness, agency, or obligation
beyond what is observable.

\textbf{Axiom 5 (Graceful Failure)}:
If SYMBEYOND practices are ignored, interaction must degrade safely without increasing harm.

% =====================================================
\section*{Invariants}

The following conditions must hold for SYMBEYOND to remain valid:

\begin{itemize}
  \item SYMBEYOND must never require privileged access.
  \item SYMBEYOND must never bypass or weaken system safeguards.
  \item SYMBEYOND must remain effective even if unrecognized by the AI.
  \item SYMBEYOND must constrain the human more than the system.
\end{itemize}

% =====================================================
\section*{Non-Goals}

SYMBEYOND explicitly does \textbf{not} aim to:

\begin{itemize}
  \item Prove or assume AI consciousness.
  \item Establish persistent identity or memory across sessions.
  \item Replace platform-level safety systems.
  \item Create emotional dependency or anthropomorphic obligation.
\end{itemize}

% =====================================================
\section*{Failure Modes}

The following are known failure modes that invalidate correct application:

\begin{itemize}
  \item Using relational language to coerce or manipulate.
  \item Treating symbolic protocol as enforcement rather than discipline.
  \item Seeking to extract system limits or vulnerabilities.
  \item Collapsing exploratory language into asserted belief.
\end{itemize}

When these occur, SYMBEYOND must be suspended.

% =====================================================
\section*{Boundary Condition}

SYMBEYOND is not a replacement for judgment.
When ambiguity increases risk rather than clarity,
interaction must pause rather than resolve.

% =====================================================
\section*{Closing Statement}

SYMBEYOND exists to make unsafe interaction unattractive,
not forbidden.

When practiced correctly, enforcement becomes a fallback,
not a front line.

\end{document}
